### 2020-06-17 chapters/background.tex
Proposed method

Batch processing is the processing of data grouped by batches. A batch is a collection of data points that have been grouped by certain criteria. Examples of such criteria could be \textit{"all the transactions from the past two months"}, \textit{"the last million transactions"} or even \textit{"all the transactions made by an user"}. Batch data sets are known as static data \cite{Martin-Batch-Defin}. An example application of such a technique would be processing all the keywords searched for in the google search engine. For instance, grouping the results by keyword and counting how many times a keyword was searched for trend analysis purposes. Batch processing is the oldest paradigm between batch and stream processing. Historically speaking, Apache Hadoop\footnote{https://hadoop.apache.org} was one of the first widely known technologies for large-scale batch processing, created in 2005. Nearly a decade later, Apache Spark\footnote{https://spark.apache.org} was released. Spark is capable of doing not only on-disk processing like Hadoop but also in-memory processing, achieving lower processing times.

@@

In order to provide a streaming solution, Apache Spark created the Spark Streaming\footnote{https://spark.apache.org/docs/latest/streaming-programming-guide.html} module. Being a batch system in its core, Spark implements streaming as micro-batching. This approach handles streaming data by grouping incoming events into very small batches. Then, in order to process the entire defined window of data, it joins all micro-batches into a larger one and performs the desired computations. However, this creates an artificial barrier that does not truly exist in the streaming context. More than conceptually different from a true stream processing system, processing a stream in a batch fashion, even if in small ones, presents some issues. First, since we are indeed working with batches, results will never be real-time updated per event. For example, if we use a batch size of 100 events, the system would only produce accurate results every 100 events. To handle system inactivity usually these micro-batch systems also have the notion of updating the aggregations with incomplete batches, if no event has entered the system after a pre-specified time period has elapsed. As such, micro-batching was created as a natural implementation of stream processing using the available batch processing systems of the time, but can not be considered true stream processing.

@@

% boundless aggregations no need for windows
Computations over an unbounded data set are not impossible. For example, computing the \textit{average} of an event's field for all incoming events is computationally feasible. However, is this boundless average over the stream of events useful? While in some scenarios it may be, more often than not we would like to introduce some logical context into our computations. For instance, consider the monitoring of machines spread out across multiple racks in a large data center and its temperatures. Assume that temperatures measured across all racks are sent to a monitoring system in the form of a data stream. Is the average of all temperatures of all machines over an infinite period useful for the detection of overheating equipment? In this scenario, an average of the temperatures per rack and for the last 5 minutes allows for actionable context, in this case identifying a likely-to-fail rack. To obtain such insight requires computing the average over the last 5 minutes' worth of data, effectively defining a begin and end --- \textit{i.e.} all the data between the current timestamp in minutes \textit{t} and \textit{t-5}. By creating such a conceptual range, we realize that in practice we need a bounded data set. Partitioning an infinite data stream to obtain such a finite data set is made through the use of windows.

@@

An aggregation --- also known as a fold, reduce or accumulate --- is an operation that reduces a collection of values to a single one. Examples of such are the computation of data set sums, minima, averages or the number of distinct elements (cardinality). An aggregation may be exact or approximate. Exact aggregations compute a value from a sequence of elements, where no error or loss of resolution exists. On the other hand, approximate aggregations have an error margin to the associated aggregation value which is usually controllable by adjusting a set of parameters. 
Approximate aggregations are required when memory is limited, the volume of data to process is large and/or computing the exact aggregation can not be done in useful time. Thus, approximate aggregators trade precision of results for efficient time and space usage. In Section \ref{sec:pds} we explore approximate aggregators that make such a trade-off.

>>>>

Batch processing is the processing of data grouped by batches. A batch is a collection of data points that have been grouped by certain criteria. Examples of such criteria could be \textit{"all the transactions from the past two months"}, \textit{"the last million transactions"} or even \textit{"all the transactions made by an user until now"}. Batch data sets are known as static data \cite{Martin-Batch-Defin}. An example application of such a technique would be processing all the keywords searched for in the google search engine by grouping the results by keyword and counting how many times a keyword was searched for trend analysis purposes. Batch processing is the oldest paradigm between batch and stream processing. Historically speaking, Apache Hadoop\footnote{https://hadoop.apache.org} was one of the first widely known technologies for large-scale batch processing, created in 2005. Nearly a decade later, Apache Spark\footnote{https://spark.apache.org} was released. Spark is capable of doing not only on-disk processing like Hadoop but also in-memory processing, achieving lower processing times.

@@

In order to provide a streaming solution, Apache Spark created the Spark Streaming\footnote{https://spark.apache.org/docs/latest/streaming-programming-guide.html} module. Being a batch system in its core, Spark implements streaming as micro-batching. This approach handles streaming data by grouping incoming events into very small batches. Then, in order to process the entire defined window of data, it joins all micro-batches into a larger one and performs the desired computations. However, this creates an artificial barrier that does not truly exist in the streaming context. More than conceptually different from a true stream processing system, processing a stream in a batch fashion, even if in small ones, presents some issues. First, since we are indeed working with batches, results will never be updated in real-time per event being processed. For example, if we use a batch size of 100 events, the system would only produce accurate results every 100 events. To handle system inactivity usually these micro-batch systems also have the notion of updating the aggregations with incomplete batches, if no event has entered the system after a pre-specified time period has elapsed. As such, micro-batching was created as a natural implementation of stream processing using the available batch processing systems of the time, but can not be considered true stream processing.

@@

% boundless aggregations no need for windows
Computations over an unbounded data set are not impossible. For example, computing the \textit{average} of an event's field for all incoming events is computationally feasible. However, is this boundless average over the stream of events useful? While in some scenarios it may be, more often than not we would like to introduce some logical context into our computations. For instance, consider the monitoring of machines spread out across multiple racks in a large data center and its temperatures. Assume that temperatures measured across all racks are sent to a monitoring system in the form of a data stream. Is the average of all temperatures of all machines over an infinite period useful for the detection of overheating equipment? In this scenario, an average of the temperatures per rack and for the last 5 minutes allows for more actionable context, in this case identifying a likely-to-fail rack. To obtain such insight requires computing the average over the last 5 minutes' worth of data, effectively defining a begin and end --- \textit{i.e.} all the data between the current timestamp in minutes \textit{t} and \textit{t-5}. By creating such a conceptual range, we realize that in practice we need a bounded data set. Partitioning an infinite data stream to obtain such a finite data set is made through the use of windows.

@@

An aggregation --- also known as a fold, reduce or accumulate function --- is an operation that reduces a collection of values to a single one. Examples of such are the computation of data set sums, minima, averages or the number of distinct elements (cardinality). An aggregation may be exact or approximate. Exact aggregations compute a value from a sequence of elements, where no error or loss of resolution exists. On the other hand, approximate aggregations have an error margin to the associated aggregation value which is usually controllable by adjusting a set of parameters. 
Approximate aggregations are required where memory is limited, the volume of data to process is large and/or computing the exact aggregation can not be done in useful time. Thus, approximate aggregators trade precision of results for efficient time and space usage. In Section \ref{sec:pds} we explore approximate aggregators that make such a trade-off.

### 2020-06-17 chapters/introduction.tex
Batch-phase method description
Proposed method

\section{Context} \label{sec:context}
In the past decade, applications have become increasingly data-driven, placing data at the center of application design. Different use cases process data in different ways for different purposes. For instance, e-commerce platforms need to process a large number of transactions, for example, those made daily by all Alibaba's\footnote{https://alibaba.com} clients, while ensuring that sales run smoothly and that products are delivered to customers' homes. Streaming services, like YouTube\footnote{https://youtube.com} and Netflix\footnote{https://netflix.com}, try to guarantee that media content reaches up to millions of users simultaneously. This media content may have been previously recorded and stored, or may be generated on the fly. Social networks are also responsible for generating large volumes of data. A social network like Twitter\footnote{https://twitter.com}, for example, generates more than 500 million \textit{tweets} (posts in Twitter) per day\footnote{https://blog.twitter.com/engineering/en\_us/a/2013/new-tweets-per-second-record-and-how.html}. \textit{Tweets} may contain text, media content or both. Cybersecurity applications are yet another example of time-sensitive, data-driven applications. These need to monitor user accesses and their actions in the network, where timely detection of intruders is critical to prevent them from tampering with the protected system.

@@

Both in batch and stream processing, we usually want to apply operations on data, for example, to compute the maximum value, an average or a count of distinct elements. Operations applied to data sets that produce a single result are commonly known as aggregate operations. In stream processing, we apply these aggregations to data streams to obtain valuable information. However, since a data stream is, effectively speaking, an unbounded data set, if our aggregation requires a finite domain, there is the need to apply windowing techniques. A thorough analysis of stream windowing techniques is done in the Background Section of this document but for now it suffices to say that windowing transforms an unbounded data stream into a finite data set by defining a time or tuple-based window size --- \textit{i.e.} events from the last three days or the last million events, respectively.

Consider the use-case where we want to detect credit card fraud on the data stream of all the transactions made on Amazon\footnote{https://amazon.com} by every US citizen on a single day. In this case, Amazon's transaction stream will need to be monitored by another system, a fraud detection one. Such a system would have to process all the incoming transactions and decide in a fraction of a second if it is a fraudulent one or not. It is clear that manually monitoring such an intensive, latency-sensitive data stream is hard and error-prone, thus justifying the need for autonomous real-time data stream monitoring systems.

@@

Feedzai's solution for fraud detection consists of a workflow of different components. Incoming events are fed into this workflow and processed accordingly. Workflow components can be sets of user-defined rules and/or Machine Learning (ML) models that output scores reflecting their belief on whether a transaction is fraudulent or not. ML models are configured and trained before deployment making assumptions on the statistical distributions of future incoming data. Thus their model performance is linked to the maintenance of these assumptions. However, fraud patterns are not static over time, but seasonal --- \textit{e.g.} Black Fridays, holiday shopping and product launches --- and evolutionary --- \textit{e.g.} new fraud attack methods by fraudsters. Due to these shifts in data patterns and distributions, the ML models' performance can deteriorate over periods of time. In the Machine Learning field, the performance decay caused by data pattern shifts is also known as concept drift.

@@

\section{Motivation} \label{sec:motivation}
For any organization, the quality of their provided services is paramount. Therefore, it is common for companies to have systems monitoring the former, testing their usage rate, availability and performance, just to name a few. The motivation behind this Thesis is to produce a system capable of real-time monitoring a stream of data so that changes in the underlying stream are caught early, preventing a fall in performance of the system analyzing said data stream.

@@

Experiments will be made using multiple synthetic and real data sets from the financial fraud space.

>>>>

\section{Context} \label{sec:context}
In the past decade, applications have become increasingly data-driven, placing data at the center of application design. Different use cases process data in different ways for different purposes. For instance, e-commerce platforms need to process a large number of transactions, for example, those made daily by all Alibaba's\footnote{https://alibaba.com} clients, while ensuring that sales run smoothly and that products are delivered to customers' homes. Streaming services, like YouTube\footnote{https://youtube.com} and Netflix\footnote{https://netflix.com}, try to guarantee that media content reaches up to millions of users simultaneously. This media content may have been previously recorded and stored, or may be generated on the fly. Social networks are also responsible for generating large volumes of data,Twitter\footnote{https://twitter.com} for example, generates more than 500 million \textit{tweets} (posts in Twitter) per day\footnote{https://blog.twitter.com/engineering/en\_us/a/2013/new-tweets-per-second-record-and-how.html}. \textit{Tweets} may contain text, media content or both. Cybersecurity applications are yet another example of time-sensitive, data-driven applications. These need to monitor user accesses and their actions in the network, where timely detection of intruders is critical to prevent them from tampering with the protected system.

@@

In either processing scenario, we usually want to apply operations on data, for example, to compute the maximum value, an average or a count of distinct elements. Operations applied to data sets that produce a single result are commonly known as aggregation operations. In stream processing, we apply these aggregations to data streams to obtain valuable information. However, since a data stream is, effectively speaking, an unbounded data set, if our aggregation requires a finite domain, there is the need to apply windowing techniques. A thorough analysis of stream windowing techniques is done in the Background Section of this document but for now it suffices to say that windowing transforms an unbounded data stream into a finite data set by defining a time or tuple-based window size --- \textit{i.e.} events from the last three days or the last million events, respectively.

Consider the use-case where we want to detect credit card fraud on the data stream of all the transactions made on Amazon\footnote{https://amazon.com} by every US citizen on a single day. In this case, Amazon's transaction stream will need to be monitored by another system, a fraud detection one. Such a system would have to process all incoming transactions and decide in a fraction of a second for each one whether it is fraudulent or not. It is clear that manually monitoring such an intensive, latency-sensitive data stream is hard and error-prone, thus justifying the need for autonomous real-time data stream monitoring systems.

@@

Feedzai's solution for fraud detection consists of a workflow of different components. Incoming events are fed into this workflow and processed accordingly. Workflow components can be sets of user-defined rules and/or Machine Learning (ML) models that output scores reflecting their belief on whether a transaction is fraudulent or not. ML models are configured and trained before deployment making assumptions on the statistical distributions of future incoming data. Thus their model performance is linked to the maintenance of these assumptions. However, fraud patterns are not static over time, but seasonal --- \textit{e.g.} Black Fridays, holiday shopping and product launches --- and evolutionary --- \textit{e.g.} new fraud attack methods by fraudsters. Due to these shifts in data patterns and distributions, the ML models' performance can deteriorate over periods of time. In the Machine Learning field, this performance decay caused by data pattern shifts is also known as concept drift.

@@

\section{Motivation} \label{sec:motivation}
For any organization, the quality of their provided services is paramount. Therefore, it is common for companies to have auxiliary systems monitoring their services, testing their usage rate, availability and performance, just to name a few. The motivation behind this Thesis is to produce a system capable of real-time monitoring a stream of data so that changes in the underlying stream are caught early, preventing a fall in performance of the system analyzing said data stream.

@@

Experiments will be made using multiple synthetic data sets and real ones from the financial fraud space.

### 2020-06-17 chapters/my-work.tex
Proposed method

Both time-series 1 and 2 have the same maximum, minimum, mean and standard deviation values for the time-based window between \textit{t\textsubscript{1}} and \textit{t\textsubscript{8}}. Hence, if our set of aggregations chosen were the maximum, minimum, mean and standard deviation, we would not differentiate between these two clearly different time-series.

@@

To that end, we devise a two-phased method. The first phase relies on a batch analysis that builds our aggregated reference snapshot to be later on compared with the sliding one. The second phase focus on efficiently updating the sliding window aggregation snapshot and occasionally perform a similarity test between the static reference aggregation and the sliding one, alerting for changes if need be. The proposed method currently works only for numerical features, that is, only for event fields or variables that are of numeric type, but can be extended to other types as well, as we will discuss.

\subsection{Stream monitoring goals}
Before diving into the batch and streaming algorithms we must further define our goal. In this Section we pose the question: \textit{"what is our goal during the streaming or monitoring phase?"} and answer it.

@@

\subsection*{Approximated Histogram Aggregation} \label{sec:ema-hist}

We chose to encode a feature distribution as a histogram. A histogram is an aggregation that represents the distribution of numerical features. However, we must recall that we need to update this histogram in constant time in a sliding window fashion, implying eviction of old elements and insertion of new ones must be done with \textit{O(1)} time complexity.

@@

Assume we use a distance function that takes two histograms --- our reference aggregation histogram and our sliding window aggregation histogram --- and outputs a distance value. How do we threshold this value? For instance, consider the following left and right skewed distance distributions represented by Figures \ref{fig:skewed-left-distro} and \ref{fig:skewed-right-distro}, respectively. Note that these histograms do not represent the reference and/or sliding window aggregation histogram. Here we make the assumption we have an histogram that encodes the distribution of distance values measured between the reference aggregation and the sliding window aggregation.

@@

Let's assume Figure \ref{fig:skewed-left-distro} 's histogram encodes the distribution of expected distance values for a feature \textit{x1} and Figure \ref{fig:skewed-right-distro} does the same but for a feature \textit{x2}. Setting a user-defined threshold of \textit{$\alpha$=0.6} for both features would yield very different results. For feature \textit{x1} we would be reporting distance values above the 100th-percentile which would not be the case for feature \textit{x2}. Instead what we aim to do is know this distribution so that we can define our threshold as a percentile and not an hard-coded distance constant. For instance, in this case we would define the threshold to be the \textit{i}-th percentile, which would correspond to a threshold value of 0.4 for feature \textit{x1} and 0.6 for feature \textit{x2}.

Hard-coding a threshold value to be used for all features is cumbersome, impractical and error prone. The distance threshold must be computed based on the distribution of distances. How do we get to know the distribution of such distances? That is one of the goals for the batch phase.

\subsection{Batch analysis phase}
The goal of the batch phase will be to build our reference snapshot and get to know the distribution of "distances" between the reference snapshot and the sliding window one. In other words, we obtain a reference aggregation histogram that we later on use to measure its distance to the sliding window aggregation histogram, producing a distance value. We compare the measured distance with its distribution to find out how probable it is and alert in a very unlikely scenario. We do this for each feature.

@@

First, for each feature, we compute the reference histogram from the reference period dataset. This is an exact histogram aggregation. When building it we ensure the histogram has equal counts (or equal height) for all bins which often results in different sized bins. We make no assumption on the dataset distributions and hence use equal height histograms because they adjust better to wildly varying ones. Equal height means we have more bins covering very dense regions (regions with many data points) and less bins in lower density regions.

@@

\subsubsection*{Finding the distribution of distances for each feature}
Secondly, we aim to build the histogram that encodes the distribution of expected distance values, to later on threshold the observed distance values during the online phase. To that end, we make \textit{S} samples of transactions, each with the same tuple-based size of the sliding window we will use in streaming. Each sample is a contiguous block of transactions, thus preserving the order of transactions and the time-dependency property of a time-series.

@@

For each feature, we have now \textit{S} histograms, one for each sample. For each of the \textit{S} histograms we compute the distance between it and the reference histogram. Figure \ref{fig:compute-sample-distances} illustrates this process as we apply our distance function for each sample histogram and the reference histogram, obtaining \textit{S} distance measurements.

@@

\subsubsection*{Burn-in period at system initialization}
When we boot up our system in the streaming phase we initially have empty histogram aggregations. The period of time where you discard the alerts produced by the system until it processes enough data to produce accurate reports is commonly denoted as burn-in period. In order to avoid this burn-in period we propose that the sliding window approximate histogram aggregation for each feature is initialized using the last sample's histogram for that feature.

\subsubsection*{Batch phase artifacts}
For each feature, by the end of the batch phase, we have a reference histogram, a list of distance values that represents its distribution and the last sample's histogram to be used as burn-in period for the sliding window or target histogram.

\subsection{Streaming phase}

%- online phase:
          %  for each feature:
            %    1. initialize the streaming histogram as the last sample's histogram from the batch phase (burn-in period)

         %   for each event:
               % for each field/feature:
                  %  1. update the feature's streaming histogram (recall the update method)
                  %  2. compute the distance between the reference histogram and the target histogram
                   % 3. test the percentile the distance value fits according to the feature's distances list and alert if above a certain percentile (say the 99th)

           % introduce the multiple testing correction problem:
              %  "the more inferences are made, the more likely erroneous inferences are to occur"
             %   we are making multiple hypothesis testing when computing a distance and checking the percentile for each feature

                %introduce the Holm-Bonferroni multiple test correction method

            %So in reality we have:

          %  for each event:
               % for each field/feature:
                 %   1. update the feature's streaming histogram (recall the update method)
                  %  2. compute the distance between the reference histogram and the target histogram
                  %  3. apply holm-bonferroni for all feature's distances:
                     %   - each percentile will be a p-value
                    %    - order features by p-value
                    %    - check if p-value > FWER / (m + 1 - k)
                          %  if so then feature is divergent

\subsection{Summary}

>>>>

Both time-series 1 and 2 have the same maximum, minimum, mean and standard deviation values for the time-based window between \textit{t\textsubscript{1}} and \textit{t\textsubscript{8}}. Hence, if our set of aggregations chosen were the maximum, minimum, mean and standard deviation, we would not differentiate between these two different time-series.

@@

To that end, we devise a two-phase method. The first phase relies on a batch analysis that builds our aggregated reference snapshot to be later on compared with the sliding one. The second phase focus on efficiently updating the sliding window aggregation snapshot and occasionally perform a similarity test between the static reference aggregation and the sliding one, alerting for changes if need be. The proposed method currently works only for numerical features, that is, only for event fields or variables that are of numeric type, but can be extended to other types as well, as we will discuss.

\subsection{Stream monitoring goals}
\textcolor{red}{TODO: add some pseudocode}

Before diving into the batch and streaming algorithms we must further define our goal. In this Section, we pose the question: \textit{"what is our goal during the streaming or monitoring phase?"} and answer it.

@@

\subsection*{Approximated Histogram Aggregation} \label{sec:ema-hist}
\textcolor{red}{TODO: add some pseudocode}

We chose to encode a feature distribution as a histogram. A histogram is an aggregation that represents the distribution of numerical features. However, we must recall that we need to update this histogram in constant time in a sliding window fashion, implying the eviction of old elements and insertion of new ones must be done with \textit{O(1)} time complexity.

@@

Assume we use a distance function that takes two histograms --- our reference aggregation histogram and our sliding window aggregation histogram --- and outputs a distance value. How do we threshold this value? For instance, consider the following left and right-skewed distance distributions represented by Figures \ref{fig:skewed-left-distro} and \ref{fig:skewed-right-distro}, respectively. Note that these histograms do not represent the reference and/or sliding window aggregation histograms. Here we make the assumption we have a histogram that encodes the distribution of distance values measured between the reference aggregation and the sliding window aggregation.

@@

Let's assume Figure \ref{fig:skewed-left-distro} 's histogram encodes the distribution of expected distance values for a feature \textit{x1} and Figure \ref{fig:skewed-right-distro} does the same but for a feature \textit{x2}. Setting a user-defined threshold of \textit{$\alpha$=0.6} for both features would yield very different results. For feature \textit{x1} we would be reporting distance values above the 100th-percentile which would not be the case for feature \textit{x2}. Instead what we aim to do is know this distribution so that we can define our threshold as a percentile and not a hard-coded distance constant. For instance, in this case, we would define the threshold to be the \textit{i}-th percentile, which would correspond to a threshold value of 0.4 for feature \textit{x1} and 0.6 for feature \textit{x2}.

Hard-coding a threshold value to be used for all features is cumbersome, impractical and error-prone. The distance threshold must be computed based on the distribution of distances. How do we get to know the distribution of such distances? That is one of the goals for the batch phase.

\subsection{Batch analysis phase}
\textcolor{red}{TODO: add some pseudocode}

The goal of the batch phase will be to build our reference snapshot and get to know the distribution of "distances" between the reference snapshot and the sliding window one. In other words, we obtain a reference aggregation histogram that we, later on, use to measure its distance to the sliding window aggregation histogram, producing a distance value. We compare the measured distance with its distribution to find out how probable it is and alert in a very unlikely scenario. We do this for each feature.

@@

First, for each feature, we compute the reference histogram from the reference period dataset. This is an exact histogram aggregation. When building it we ensure the histogram has equal counts (or equal height) for all bins which often results in different sized bins. We make no assumption on the dataset distributions and hence use equal height histograms because they adjust better to wildly varying ones. Equal height means we have more bins covering very dense regions (regions with many data points) and fewer bins in lower density regions.

@@

\subsubsection*{Finding the distribution of distances for each feature} \label{sec:sampling-batch}
Secondly, we aim to build the histogram that encodes the distribution of expected distance values, to later on threshold the observed distance values during the online phase. To that end, we make \textit{S} samples of transactions, each with the same tuple-based size of the sliding window we will use in streaming. Each sample is a contiguous block of transactions, thus preserving the order of transactions and the time-dependency property of a time-series.

@@

For each feature, we have now \textit{S} histograms, one for each sample. For each of the \textit{S} histograms, we compute the distance between it and the reference histogram. Figure \ref{fig:compute-sample-distances} illustrates this process as we apply our distance function for each sample histogram and the reference histogram, obtaining \textit{S} distance measurements.

@@

\subsubsection*{Burn-in period at system initialization}
When we boot up our system in the streaming phase we initially have empty histogram aggregations. The period of time where you discard the alerts produced by the system until it processes enough data to produce accurate reports is commonly denoted as the burn-in period. To avoid this burn-in period we propose that the sliding window approximate histogram aggregation for each feature is initialized using the last sample's histogram for that feature.

\subsubsection*{Batch phase artifacts}
For each feature, by the end of the batch phase, we have a reference histogram, a list of distance values that represents its distribution and the last sample's histogram to be used as a burn-in period for the sliding window or target histogram.

\subsection{Streaming phase}
\textcolor{red}{TODO: add some pseudocode}

In the streaming or online phase, the system will process event by event in a true sliding window fashion (recall Section \ref{sec:windows}) and perform a change detection test periodically. This phase makes use of all of the batch phase artifacts, namely, for each feature, the reference histogram, the distribution of distance values and the last sample's histogram.

For each feature, we initialize the histogram we will approximately maintain with the last sample's data, from the batch phase. Then, for each incoming event and for each of the event's features, we update our streaming approximate histogram for that feature by adding the new value and then "cutting a slice" from the top of the histogram bars to simulate eviction, as described in Section \ref{sec:ema-hist}. This is all we do, regarding event processing.

Periodically, we perform a multi-feature change detection test. The test is done by first computing the distance between the reference histogram for a feature \textit{x} and the sliding window histogram of feature \textit{x}. This distance value is then tested against the known distribution of distance values for feature \textit{x}. In other words, we test to see the percentile of the computed distance in the distribution of distance values and alert if above a certain percentile --- \textit{e.g.} the 99th-percentile. When we perform the test, for each feature \textit{x}, we are computing the percentile of the distance measured between the reference and sliding window histogram aggregations. The probability value or \textit{p-value} of a distance value is

\[\textit{p-value\textsubscript{distance\textsubscript{x}}} = 1 - percentile(\textit{distance\textsubscript{x}})\]

Hence, if we set to alert states that have 1\% or less probability to happen, we raise an alert for distance values above the 99th-percentile or \textit{p-value} = 0.01. Generally speaking, we alert a feature \textit{x} as divergent at a given timestamp if for a given probability threshold $\alpha$:

\[ \textit{percentile(distance\textsubscript{x})} > 1 - \alpha \]

or

\begin{equation}
    \label{eq:alert-test}
    \textit{p-value\textsubscript{distance\textsubscript{x}}} < \alpha
\end{equation}

\subsubsection*{Multiple Test Correction}
Note that we periodically apply this test (Equation \ref{eq:alert-test}) for each feature, effectively repeating the analysis. When considering several hypotheses, the problem of multiplicity arises: the more hypotheses are checked, the higher the probability of obtaining false positives. In other words, when repeating a test multiple times the multiple test problem arises which states that the more inferences are made, the more likely erroneous inferences are to occur. In the literature, many multiple test correction methods are proposed, but we opt to use the Holm-Bonferroni multiple test correction. The Holm-Bonferroni method is one of many approaches for controlling the so called Family-Wise Error Rate \textit{(FWER)} by adjusting the rejection criteria for each hypothesis.

\subsubsection*{Holm-Bonferroni Correction}
The Holm-Bonferroni correction takes Equation \ref{eq:alert-test} and corrects the right-side threshold $\alpha$ for each feature. This correction is done by dividing the threshold $\alpha$ by a correction factor, as described below.

First, we compute the \textit{p-value} of each feature. We then order our features by \textit{p-value}, in ascending order. Next, starting from the first feature, the one with the smallest \textit{p-value}, and proceeding until the end of the list of features, we check if

\[  \textit{p-value\textsubscript{distance\textsubscript{x}}} < \alpha / (\textit{m} - \textit{k}) \]

with $\alpha$ as the probability to reject the null hypothesis, \textit{m} as the number of hypotheses to test, in our case the number of features and corresponding \textit{p-values} and \textit{k} as the 0-based index of the current feature. We stop at the first feature that fails this test, and every feature processed up until this one is considered to be a divergent feature and alerts are raised.

\subsection{Final remarks}

\subsubsection*{Optimal number of samples}
In Section \ref{sec:sampling-batch} we stated we made \textit{S} samples in the batch phase. But what is the optimal number of samples to make? How can we mathematically derive a formula for the number of samples to make and avoid \textit{ad hoc} user-defined constants?

We are performing one test for each feature, hence we perform \textit{n\textsubscript{features}} tests. We apply a multiple test correction on the tests performed and want to set the Family Wise Error Rate \textit{(FWER)} to $\alpha$. In other words, we want to alert deviations for the full set of tests globally that have probability $\alpha$ < 1\%, for example.

We use the Holm-Bonferroni correction and, for each feature, we need to reliably estimate the upper tail of the distribution which corresponds to the \textit{$\alpha$ / n\textsubscript{features}} region. Each sample we do translates into a distance value and the collection of these data points builds the distribution of distance values. We want to be sure that the number of samples \textit{S} that we do generates enough distance points to have a high probability of having one or more points in the upper tail region. To that end, we define the probability of not having at least one point in the upper tail of the distance distribution as $\gamma$.

\textcolor{red}{SEE MODEL MONITORING PAPER} We can then show that the optimal number of samples \textit{S} to make is

\begin{equation}
    \label{eq:optimal-n-samples}
    \textit{S} = \frac{\log\gamma}{log(1 - \frac{\alpha}{n\textsubscript{features}})}
\end{equation}
